---
title: "creating_stim_sets_while_running_experiment"
author: "Christian Stenbro"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
pacman::p_load(tidyverse,
               rjson, 
               jsonlite)
```

Importing the original stim set files:

```{r}
study_1_stim <- fromJSON("/Users/christianstenbro/Programming/eSYMb_similarity_git/eSYMb_similarity_rating_project/_static/similarity/stim_lists/new_stim_list_study_1_5_ratings.JSON")
study_2_stim <- fromJSON("/Users/christianstenbro/Programming/eSYMb_similarity_git/eSYMb_similarity_rating_project/_static/similarity/stim_lists/new_stim_list_study_2_5_ratings.JSON")
study_3_stim <- fromJSON("/Users/christianstenbro/Programming/eSYMb_similarity_git/eSYMb_similarity_rating_project/_static/similarity/stim_lists/new_stim_list_study_3_5_ratings.JSON")
```

We want to do the following: 

- Split the stim sets into two separate objects: One containing the already rated sets (object 1), one with the remaining sets (object 2)
- Object 1 should be stored in a new, continuously updated set called 'already_rated_sets'
- Object 2 should be stored in a new, continuously updated set called 'unrated_sets'. It is important that this list is reformatted to that it starts from key "1".

This needs to be done for all three studies. The unrated sets are then imported into Otree (possibly using the same file name as the ones in there now to avoid having to rename variables across all the pages).

To do this, we need to keep the number of participants that have already completed each study in mind: 

```{r}
completed_participants_study_1 <- 3
completed_participants_study_2 <- 3
completed_participants_study_3 <- 4
```

These can be used as parameters in the following function:

```{r}
create_new_sets <- function(study_stim_set, num_completed_participants) {
  
  n = num_completed_participants

  set_size = length(study_stim_set)
  
  already_rated <- study_stim_set[1:n]
  
  unrated <- study_stim_set[(n+1):set_size]
  
  # making new names starting from one:
  new_names <- 1:(set_size-n)
  
  # assigning new names to the unrated set:
  names(unrated) <- new_names
  
  # this then needs to be stored as a unique object
  return(list(unrated_sets = unrated,
              already_rated_sets = already_rated))
}
```

Running the function on sets from all studies:

```{r}
study_1_new_sets <- create_new_sets(study_stim_set = study_1_stim,
                                    num_completed_participants = completed_participants_study_1)

study_2_new_sets <- create_new_sets(study_stim_set = study_2_stim,
                                    num_completed_participants = completed_participants_study_2)

study_3_new_sets <- create_new_sets(study_stim_set = study_3_stim,
                                    num_completed_participants = completed_participants_study_3)

# we can check that the 1st entry of a new sets match the nth entry of an old set
for (i in 1:length(study_1_new_sets$unrated_sets)) {
  n = completed_participants_study_1
  print(unique(study_1_new_sets$unrated_sets[[i]] == study_1_stim[[i+n]]))
}

for (i in 1:length(study_2_new_sets$unrated_sets)) {
  n = completed_participants_study_2
  print(unique(study_2_new_sets$unrated_sets[[i]] == study_2_stim[[i+n]]))
}

for (i in 1:length(study_3_new_sets$unrated_sets)) {
  n = completed_participants_study_3
  print(unique(study_3_new_sets$unrated_sets[[i]] == study_3_stim[[i+n]]))
}
```

```{r}
# checking that the length matches the expectations
length(study_1_new_sets$unrated_sets)
length(study_2_new_sets$unrated_sets)
length(study_3_new_sets$unrated_sets)
```

Everything looks okay so we move on to storing the sets:

```{r}
# storing the unrated sets individually
updated_stim_set_study_1 <- study_1_new_sets$unrated_sets
updated_stim_set_study_2 <- study_2_new_sets$unrated_sets
updated_stim_set_study_3 <- study_3_new_sets$unrated_sets

# checking that the names are correct
names(updated_stim_set_study_1)
names(updated_stim_set_study_2)
names(updated_stim_set_study_3)
```

```{r}
# converting back to JSON format
updated_stim_set_study_1_JSON <- jsonlite::toJSON(updated_stim_set_study_1, pretty = TRUE)
updated_stim_set_study_2_JSON <- jsonlite::toJSON(updated_stim_set_study_2, pretty = TRUE)
updated_stim_set_study_3_JSON <- jsonlite::toJSON(updated_stim_set_study_3, pretty = TRUE)
```

```{r}
# function to save as JSON file
finalize_set <- function(JSON_set_list, path, file_name, study_nr, number_of_ratings) {
  
  iteration = paste0("_study_", study_nr, "_", number_of_ratings, "_ratings")
  
  extension = ".JSON"
  
  write(JSON_set_list, file = paste0(path, file_name, iteration, extension))

}
```

Saving sets - notice that the file names are identical to the complete sets. The idea is to drag these into the stim_lists folder in the static folder and overwrite the old files.

```{r}
# saving new set for study 1
finalize_set(JSON_set_list = updated_stim_set_study_1_JSON,
             path = "/Users/christianstenbro/AU/eSYMb/Similarity Rating Exp/Pilot/updated_stim_sets/",
             file_name = "new_stim_list",
             study_nr = 1,
             number_of_ratings = 5)

# saving new set for study 2
finalize_set(JSON_set_list = updated_stim_set_study_2_JSON,
             path = "/Users/christianstenbro/AU/eSYMb/Similarity Rating Exp/Pilot/updated_stim_sets/",
             file_name = "new_stim_list",
             study_nr = 2,
             number_of_ratings = 5)

# saving new set for study 3
finalize_set(JSON_set_list = updated_stim_set_study_3_JSON,
             path = "/Users/christianstenbro/AU/eSYMb/Similarity Rating Exp/Pilot/updated_stim_sets/",
             file_name = "new_stim_list",
             study_nr = 3,
             number_of_ratings = 5)
```
