---
title: "testing similarity data"
author: "Christian Stenbro"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
pacman::p_load(tidyverse,
               rjson,
               rethinking)
```

```{r}
# reading pilot data
data <- read_csv("/Users/christianstenbro/AU/eSYMb/Similarity Rating Exp/Pilot/otree_data/Pilot_all_studies_session_3_SimilarityRatingExperiment_2025-06-19.csv")
```
# Pre-processing

```{r}
# filtering out non-participant rows
data <- data %>% filter(participant.label != "NA")

# checking out the data for a single participant (participant 1)
participant_list <- unique(data$participant.label)
```

Turn this into a function:

```{r}
organize_data <- function(ID_number) {
  
  ID <- participant_list[ID_number]
  
  subset <- data %>% filter(participant.label == ID)
  
  # setting up constants
  rounds <- c(1,3,5)
  
  # unpacking the JSON lists
  image_indices <- list()
  image_ratings <- list()
  rating_times <- list()
  is_attention_check <- list()
  #attention_check_status <- list()
  
  for (i in rounds) {
    image_indices <- append(image_indices, list(fromJSON(subset$player.imageIndices[i])))
    image_ratings <- append(image_ratings, list(fromJSON(subset$player.imageRatings[i])))
    rating_times <- append(rating_times, list(fromJSON(subset$player.ratingTimes[i])))
    is_attention_check <- append(is_attention_check, list(fromJSON(subset$player.isAttentionCheck[i])))
    #attention_check_status <- append(attention_check_status, list(fromJSON(subset$player.withinExpectedRatingRange[i])))
  }
  
  #unpacking lists
  image_indices <- unlist(image_indices)
  image_ratings <- unlist(image_ratings)
  rating_times <- unlist(rating_times)
  is_attention_check <- unlist(is_attention_check)
  #attention_check_status <- unlist(attention_check_status)
  
  # making a new data base for visualisation purposes
  subset_unpacked <- data_frame(image_indices,
                                image_ratings,
                                rating_times,
                                is_attention_check)
                                #attention_check_status)
  
  
  
  # making a single object containing the attention check status variable
  attention_check_status <- list()
  
  for (i in rounds) {
    attention_check_status <- append(attention_check_status, list(fromJSON(subset$player.withinExpectedRatingRange[i])))
  }
  
  attention_check_status <- unlist(unlist(attention_check_status))
  
  return(list(subset_unpacked = subset_unpacked, 
              attention_check_status = attention_check_status)
         )
}
```

Making a function to append attention checks to the participant subset (relies on the previous function):

```{r}
subset_with_attention_check_status <- function(ID_number) {
  
  temp_data <- organize_data(ID_number)
  
  # getting the indices for the att. checks
  att_checks_indices <- temp_data$subset_unpacked$image_indices[temp_data$subset_unpacked$is_attention_check == TRUE]
  
  # cleaning the attention check pass fail variable to only contain TRUE / FALSE
  attention_check_status_clean <- grep("^(TRUE|FALSE)$", temp_data$attention_check_status, value = TRUE)
  
  # combining into a new table alongside the att. check indices
  attention_check_status_clean <- tibble(attention_check_status_clean, att_checks_indices)
  
  # merging this with the organized data frame
  subset_complete <- merge(temp_data$subset_unpacked, attention_check_status_clean, by.x = "image_indices", by.y = "att_checks_indices", all=T)
 
  return(subset_complete)
}
```

The final step is to stitch together a new grand data frame from the subsets created with the subset_with_attention_check_status() function:

```{r}
# setting up constants
participants <- c(1:10)

# setting up empty data frame
df <- data.frame()

# looping
for (i in participants) {
  temp_subset <- subset_with_attention_check_status(i)
  sequential_ID <- rep(i, nrow(temp_subset))
  temp_subset$sequential_ID <- sequential_ID
  df <- rbind(df, temp_subset)
}

# checking out the combined data frame
df
```

Checking attention check status

```{r}


df %>% filter(sequential_ID == 3) %>% filter(is_attention_check == TRUE)
```


# Extracting comments
```{r}
data$player.technicalIssues
data$player.dataValidity
```


# Statistics

```{r}
# grouping the data
df <- df %>% group_by(sequential_ID)

# reformating data types
df$image_ratings <- as.numeric(df$image_ratings)
df$attention_check_status_clean <- as.logical(df$attention_check_status_clean)


df_summary <- df %>% summarise('rt_median' = median(rating_times), 
                 'rt_sd' = sd(rating_times),
                 'rating_mean' = mean(image_ratings), 
                 'rating_sd' = sd(image_ratings),
                 'failed_att_checks' = sum(attention_check_status_clean == FALSE, na.rm = TRUE))

# print the entire stat summary
df_summary

# computing aggregated stats
# mean of means study 3
round(mean(df_summary$rt_median[1:4]), 1)
round(mean(df_summary$rt_sd[1:4]), 1)
round(mean(df_summary$rating_mean[1:4]), 1)
round(mean(df_summary$rating_sd[1:4]), 1)

# mean of means study 2
round(mean(df_summary$rt_median[5:7]), 1)
round(mean(df_summary$rt_sd[5:7]), 1)
round(mean(df_summary$rating_mean[5:7]), 1)
round(mean(df_summary$rating_sd[5:7]), 1)

# mean of means study 1
round(mean(df_summary$rt_median[8:10]), 1)
round(mean(df_summary$rt_sd[8:10]), 1)
round(mean(df_summary$rating_mean[8:10]), 1)
round(mean(df_summary$rating_sd[8:10]), 1)
```
The stats are appended to the data frame

```{r}
df <- merge(df, df_summary, by = "sequential_ID", all = T)
```

# Visualisation

Prior to visualising, data from participants failing more than two attention checks are filtered out:

```{r}
df_summary
```


```{r}
#visual_subset <- df %>% filter(failed_att_checks <= 2)
visual_subset <- df
```

```{r}
simplehist(visual_subset$image_ratings)
simplehist(visual_subset$rating_times)
```


```{r}
visual_subset

# visualising the combined data frame
# making color scale for ratings
z <- as.numeric(visual_subset$image_ratings)
n_colors <- 100
color_indices <- as.numeric(cut(z, breaks = n_colors, include.lowest = TRUE))
point_colors_r <- viridis(n_colors)[color_indices]

# making color scale for rating times
z <- visual_subset$rating_times
n_colors <- 100
color_indices <- as.numeric(cut(z, breaks = n_colors, include.lowest = TRUE))
point_colors_rt <- viridis(n_colors)[color_indices]

# drawing plots for image ratings
plot(x = visual_subset$image_indices,
     y = visual_subset$image_ratings,
     col = col.alpha(point_colors_r, 0.2),
     pch = 19,
     main = "ratings x indices",
     xlab = "index", ylab = "similarity rating (numeric)")

# adding stats
#abline(h = df_summary$rating_mean, col = df_summary$sequential_ID)


# drawing rating time plot
plot(x = visual_subset$image_indices,
     y = visual_subset$rating_times,
     col = col.alpha(point_colors_rt, 0.2), 
     pch = 19,
     main = "rating times x indices",
     xlab = "index", ylab = "rating time (mili seconds)",
     ylim = c(0,40000)
     )

```



```{r}
# trying to visualize this stuff
library(viridis)

# making color scale for ratings
z <- as.numeric(subset_unpacked$image_ratings)
n_colors <- 100
color_indices <- as.numeric(cut(z, breaks = n_colors, include.lowest = TRUE))
point_colors_r <- viridis(n_colors)[color_indices]

# making color scale for rating times
z <- subset_unpacked$rating_times
n_colors <- 100
color_indices <- as.numeric(cut(z, breaks = n_colors, include.lowest = TRUE))
point_colors_rt <- viridis(n_colors)[color_indices]

# drawing plots
plot(x = subset_unpacked$image_indices,
     y = subset_unpacked$image_ratings,
     col = point_colors_r,
     pch = 19,
     main = "ratings x indices",
     xlab = "index", ylab = "similarity rating (numeric)")

plot(x = subset_unpacked$image_indices,
     y = subset_unpacked$rating_times,
     col = point_colors_rt,
     pch = 19,
     main = "rating times x indices",
     xlab = "index", ylab = "rating time (mili seconds)",
     ylim = c(0,50000))
```

