---
title: "visualizing drawings"
author: "Murillo Pagnotta"
date: "2025-02-28"
output: html_document
---

# Set-up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/christianstenbro/AU/eSYMb/Similarity Rating Exp/From Murillo OneDrive/sample drawings 2")
```

```{r}
pacman::p_load(
  tidyverse,
  base64enc,
  imager,
  magick,
  ggplot2, 
  rethinking)
```

READ THE CLEAN DATA from the memorability experiment (TC 3 conditions)

```{r}
d1 = read_csv("memorability_sample_Data.csv")

colnames(d1)
# NOTE: The target drawing and the reproduction are in the following columns:
# drawing: the target drawing
# drawing_copy: the reproduction

# The drawing is encoded in base64 in this experiment, as a long string starting with "data:image/png;base64,"
d1$drawing[1]
```

# 0. Combining attention check images

```{r}
# loading attention check images into magick image objects
path = "/Users/christianstenbro/AU/eSYMb/Similarity Rating Exp/From Murillo OneDrive/attention_check_drawings"

star_path = paste0(path, "/star.png")
vertical_path = paste0(path, "/vertical.png")
xxx_path = paste0(path, "/xxx.png")

star <- image_read(path = star_path)
vertical <- image_read(path = vertical_path)
xxx = image_read(path = xxx_path)

# combining all images
sv <- image_append(c(star, vertical))
ss <- image_append(c(star, star))
vv <- image_append(c(vertical, vertical))
sx <- image_append(c(star, xxx))
xv <- image_append(c(xxx, vertical))
xx <- image_append(c(xxx, xxx))

# saving images (this should be set up as a loop . . .)
path_name = "/Users/christianstenbro/AU/eSYMb/Similarity Rating Exp/From Murillo OneDrive/attention_check_export/"
  file_name = paste0(path_name,"attention_check_","xx",".png")
  png(file_name, units="in", width=7.29, height=4.51, res=500)

plot(xx)

dev.off()
```

```{r}
# making a JSON dictionary type file for loading attention check files

# making path names
check_name <- (c("sv", "ss", "vv", "sx", "xv", "xx"))

path_names <- list()

for (i in check_name){
  path_names[i] <- print(paste0("attention_check_",i,".png"))
}

path_names <- unname(unlist(path_names))

# attention_dictionary <- data_frame(
#   "index" = c(1,2,3,4,5,6),
#   "path" = path_names)

# converting to JSON file
attention_JSON <- jsonlite::toJSON(path_names, pretty = TRUE)

# saving the JSON file
path = "/Users/christianstenbro/AU/eSYMb/Similarity Rating Exp/From Murillo OneDrive/attention_check_list/"
name = "attentionJSON"
iteration = "_1" # remember to change this to not overwrite previous sets
extension = ".JSON"

write(attention_JSON, file = paste0(path, name, iteration, extension))
```


# 1. Visualising pairs

## Visualise 10 pairs:

```{r}
 i = sample(1:nrow(d1), 1)

  target = str_split_1(d1$drawing[i], pattern = ",")[2] %>% # Remove the prefix
      base64decode() %>% # Decode base64 string to raw image data
      image_read() #%>% # Convert raw image data to a magick image object
      # image_resize("200x200")  # could use this to resize if necessary
    #plot(target)
  
  # Create a blank spacer (adjust width for more space)
  spacer <- image_blank(width = 200, height = 10, color = "white") 
  
    copy = str_split_1(d1$drawing_copy[i], pattern = ",")[2] %>% # Remove the prefix
      base64decode() %>% # Decode base64 string to raw image data
      image_read() #%>% # Convert raw image data to a magick image object
      # image_resize("200x200")  # could use this to resize if necessary
      #plot(copy)
    
    # Combine images side by side
    combined_img <- image_append(c(target, spacer, copy))  # Combine horizontally
    
    plot(combined_img)
```

```{r}

for (j in 1:10) {

  i = sample(1:nrow(d1), 1)

  target = str_split_1(d1$drawing[i], pattern = ",")[2] %>% # Remove the prefix
      base64decode() %>% # Decode base64 string to raw image data
      image_read() #%>% # Convert raw image data to a magick image object
      # image_resize("200x200")  # could use this to resize if necessary
    #plot(target)
  
  # Create a blank spacer (adjust width for more space)
  spacer <- image_blank(width = 200, height = 10, color = "white") 
  
    copy = str_split_1(d1$drawing_copy[i], pattern = ",")[2] %>% # Remove the prefix
      base64decode() %>% # Decode base64 string to raw image data
      image_read() #%>% # Convert raw image data to a magick image object
      # image_resize("200x200")  # could use this to resize if necessary
      #plot(copy)
    
    # Combine images side by side
    combined_img <- image_append(c(target, spacer, copy))  # Combine horizontally
    
    plot(combined_img)
    
}
```

## Saving 10 pairs as images

Now lets save these as image files; for example png:

```{r}
# naming batch
batch = 1

# number of images
n = 15

# setting up loop
for (j in 1:n) {

  # setting up the file attributes
  path_name = "/Users/christianstenbro/AU/eSYMb/Similarity Rating Exp/From Murillo OneDrive/drawing_export/"
  file_name = paste0(path_name,"test_pair_", j, "_batch_", batch, ".png")
  png(file_name, units="in", width=7.29, height=4.51, res=500)
  
  # sampling a random pair
  i = sample(1:nrow(d1), 1)
  
  # sampling 
  target = str_split_1(d1$drawing[i], pattern = ",")[2] %>% 
    base64decode() %>% 
    image_read() 
    
  # Create a blank spacer (adjust width for more space)
  spacer <- image_blank(width = 400, height = 10, color = "white") 
    
  copy = str_split_1(d1$drawing_copy[i], pattern = ",")[2] %>% # Remove the prefix
    base64decode() %>% # Decode base64 string to raw image data
    image_read() #%>% # Convert raw image data to a magick image object
    # image_resize("200x200")  # could use this to resize if necessary
    #plot(copy)
      
  # Combine images side by side
  combined_img <- image_append(c(target, spacer, copy))  # Combine horizontally
  plot(combined_img)
  
  # turning off device
  dev.off()
  
}
```

## Visualising all pairs:

```{r}
# checking the number of pairs
dim(d1)[1]
```

```{r}
 # naming batch
batch = 2

# number of images
n = dim(d1)[1]

# setting up loop
for (j in 1:n) {

  # setting up the file attributes
  path_name = "/Users/christianstenbro/AU/eSYMb/Similarity Rating Exp/From Murillo OneDrive/drawing_export/"
  file_name = paste0(path_name,"test_pair_", j, "_batch_", batch, ".png")
  png(file_name, units="in", width=7.29, height=4.51, res=500)
  
  # sampling a random pair
  i = sample(1:nrow(d1), 1)
  
  # sampling 
  target = str_split_1(d1$drawing[i], pattern = ",")[2] %>% 
    base64decode() %>% 
    image_read() 
    
  # Create a blank spacer (adjust width for more space)
  spacer <- image_blank(width = 400, height = 10, color = "white") 
    
  copy = str_split_1(d1$drawing_copy[i], pattern = ",")[2] %>% # Remove the prefix
    base64decode() %>% # Decode base64 string to raw image data
    image_read() #%>% # Convert raw image data to a magick image object
    # image_resize("200x200")  # could use this to resize if necessary
    #plot(copy)
      
  # Combine images side by side
  combined_img <- image_append(c(target, spacer, copy))  # Combine horizontally
  plot(combined_img)
  
  # turning off device
  dev.off()
  
}
```


# 2. Making sampling mechanism

What we really want could also be achieved in the following (perhaps slightly more elegant way):

1.	Sample 100 subsets, each with ~300 drawings from the full set

2.	The sample mechanism should ensure that:

    a.	Each subset contains no more than one copy of a unique pair (trivial)

    b.	Altogether, the subsets should be composed in a way to make sure that each pair is included in n sets; that is, is sampled n times

## Making initial division

```{r}
# defining global parameters
participant_number = 100
ratings = 3
setSize <- 1000 # normally this should be length(d1$drawing)
attention_checks <- 6
```

```{r}
# computing indices for each pair in the entire set
pairIndices <- seq(1:setSize)

# setting number of divisions (should correspond to the number of participants)
breaks <- 100

# creating ordered index
sampleIdx <- cut(pairIndices, breaks = breaks, labels = FALSE)

# scramble the index
sampleIdx <- sample(sampleIdx)

# splitting the set
smallSets <- split(pairIndices, sampleIdx)

# shuffling the order of integers within each set
smallSets <- lapply(smallSets, sample)


# figuring out if the split works
setLength <- list()

for (i in 1:length(smallSets)) {
  
  setLength[i] <- length(smallSets[[i]])
  
}

# checking that it sums up correctly; that no indices are lost
sum(unlist(setLength))

```

This would be great if we only needed a single rating pr. drawing; which is not the case though.

One solution is to combine all sets with four other sets. This can't be randomized as we need to make sure that all drawings are rated n times. 

I want to combine the sets in the following fashion:

    smallSets[[1]] + smallSets[[2]] + smallSets[[3]] + smallSets[[4]] + smallSets[[5]]
    
    smallSets[[2]] + smallSets[[3]] + smallSets[[4]] + smallSets[[5]] + smallSets[[6]]
    
    smallSets[[3]] + smallSets[[4]] + smallSets[[5]] + smallSets[[6]] + smallSets[[7]]
    
    smallSets[[4]] + smallSets[[5]] + smallSets[[6]] + smallSets[[7]] + smallSets[[8]]
    
    smallSets[[5]] + smallSets[[6]] + smallSets[[7]] + smallSets[[8]] + smallSets[[9]]
    
    .  .  .

What's the best way to do this?

We could simply make n column vectors, shifting the index by one unit for each of the five:

```{r}
# parametric version

# setting up empty list
setList <- list()

# defining the number of sets to be combined (corresponding to the number of ratings pr. drawing pair)
n <- ratings
offsets <- seq(0, n-1)

# looping through the small sets and combining them according to the specified structure
for (i in 1:100) {
  setList[[i]] <- unlist( lapply( 
    offsets, function( x ) smallSets[[( i - 1 + x ) %% 100 + 1]] # computes indices of small sets as a function of the offset parameter
    ) )
}

# randomizing order of entries within each list of drawing pair indices (to make sure no participant is presented with similar orders of drawings)
setList <- lapply(setList, sample)
```

Performing checks:

```{r}
# checking the length of the final lists of lists
length(setList)

# visualising to check that each pair appears n times across all lists
simplehist(unlist(setList))
unique(table(unlist(setList)))

# checking that none of the lists has any repeating pairs
status <- list()

for (i in 1:length(setList)) {

  status[i] <- (length(unique(setList[[i]])) == length(setList[[i]]))

  }

unique(status)
```

```{r}
unpackedSet <- tibble(unlist(setList))

colnames(unpackedSet) <- "idx"

# the length of this set should correspond to n for all indices
length(t(filter(unpackedSet, idx == "1000"))) == n
```

## Saving the setList as a JSON-file:

```{r}
# adding indices ('names') to the setList
names(setList) <- as.character(seq_along(setList))

# converting to JSON
setListJSON <- jsonlite::toJSON(setList, pretty = TRUE)

# checking the data structure
#setListJSON
```

```{r}
# saving the set
path = "/Users/christianstenbro/AU/eSYMb/Similarity Rating Exp/From Murillo OneDrive/drawing_lists/"
name = "setListJSON"
iteration = "_3" # remember to change this to not overwrite previous sets
extension = ".JSON"

write(setListJSON, file = paste0(path, name, iteration, extension))
```

I wonder whether this could be incorporated directly into the Otree experiment . . . in a sense generated live once the session is instantiated. Not sure if this is a more or less complicated solution though.

## Saving visualisations of this batch

```{r}
 # naming batch
batch = 3

# number of images
n = setSize

# setting up loop
for (j in 1:n) {

  # setting up the file attributes
  path_name = "/Users/christianstenbro/AU/eSYMb/Similarity Rating Exp/From Murillo OneDrive/drawing_export/"
  file_name = paste0(path_name,"test_pair_", j, "_batch_", batch, ".png")
  png(file_name, units="in", width=7.29, height=4.51, res=500)
  
  # sampling a random pair index
  ## comment: not sure this randomization is necessary; it is just a file list after all
  ## and the randomization is handled by the stim list, which is fully random
  i = sample(1:nrow(d1), 1, replace = FALSE)
  
  # sampling images based on randomized indices (again, not sure this is necessary)
  target = str_split_1(d1$drawing[i], pattern = ",")[2] %>% 
    base64decode() %>% 
    image_read() 
    
  # Create a blank spacer (adjust width for more space)
  spacer <- image_blank(width = 400, height = 10, color = "white") 
    
  copy = str_split_1(d1$drawing_copy[i], pattern = ",")[2] %>% # Remove the prefix
    base64decode() %>% # Decode base64 string to raw image data
    image_read() #%>% # Convert raw image data to a magick image object
    # image_resize("200x200")  # could use this to resize if necessary
    #plot(copy)
      
  # Combine images side by side
  combined_img <- image_append(c(target, spacer, copy))  # Combine horizontally
  plot(combined_img)
  
  # turning off device
  dev.off()
  
}
```